### 2024-08-22 src/14-supplemental-A.Rmd
added supplemental file to explore necessity of route-level hyperparameter

---
title: "Supplemental A - Comparison of Point-level Models with and without route-level
  hyperparameter"
author: "Brandon P.M. Edwards"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Given the structured nature of the current BBS analysis, whereby up to 50 point counts are assembled into a "route" and analyzed at the route level (by summing counts at each point), it is natural to wonder if a point-level model should consider drawing point-level effects from their own independent distributions, or if the point-level effects should be drawn from a route-level hyperparameter. This supplemental file explores the differences in indices and trends produced by point-level models with and without route-level hyperparameters. For all three of POINT, DETECT, and VARPROP, we simply compare the 1-to-1 relationship between indices and trends, respectively, for each respective model, to assess if there are any distinct patterns of bias.

```{r cars, echo = FALSE}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.